---
title: "p8105_hw1_zak2132"
author: "Zachary Katz"
date: "9/21/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

Problem 1 asks us to create a data frame comprised of:
* A random sample of size 10 from a standard normal distribution
* A logical vector indicating whether elements of the sample are greater than 0
* A character vector of length 10
* A factor vector of length 10, with 3 different factor "levels"

### Create the data frame

Let's create the data frame.

```{r create_data_frame}
rand_sample <- rnorm(10)
logical_vector <- rand_sample > 0
character_vector <- c("s", "t", "r", "a", "w", "b", "e", "r", "r", "y")
factor_vector <- factor(c("first", "second", "third", "first", "second", "third", "first", "second", "third", "first"))
test_data_frame <- data.frame(rand_sample, logical_vector, character_vector, factor_vector)
test_data_frame
str(test_data_frame)
```

#### Find the means

Now that we've created the data frame, we try to take the mean of each variable.

```{r try_mean}
sample_mean <- mean(test_data_frame$rand_sample)
mean_positive <- mean(test_data_frame$logical_vector)
mean_character_vector <- mean(test_data_frame$character_vector)
mean_factor_vector <- mean(test_data_frame$factor_vector)
```

* The mean of the random sample is `r sample_mean` and the mean of the logical vector is `r mean_positive`, which is the proportion of the random sample greater than 0.

* When we try to find the mean of the character and factor vectors, we learn that the mean function does not accept arguments that are not numeric or logical, and therefore returns NA.

#### Convert vectors

Finally, we want to convert the logical, character, and factor variables to a numeric variable if possible.

```{r convert_to_numeric, results="hide"}
convert_logical <- as.numeric(logical_vector)
convert_logical

convert_character <- as.numeric(character_vector)
convert_character

convert_factor <- as.numeric(factor_vector)
convert_factor
```
* Converting the logical vector to a numeric vector changed **true to 1** and **false to 0**. 

* Converting the factor vector to a numeric vector changed **first level to 1, second level to 2, and third level to 3**.

* Converting the character vector to a numeric vector resulted in **"NAs introduced by coercion"**, leading to a vector composed only of 10 NAs.

## Problem 2

#### Load the penguins dataset

```{r load_penguins_dataset}
data("penguins", package = "palmerpenguins")
```

#### Describe the penguins dataset

First, we want to take a quick look at data and its structure.
```{r examine_penguins_data}
penguins

str(penguins)
```

The names of the variables in the dataset are `r names(penguins)` and the size of the dataset is `r nrow(penguins)` rows and `r ncol(penguins)` columns.

We are also asked to calculate the mean flipper length, which is `r mean(penguins$flipper_length_mm, na.rm = T)` mm, after removing NA entries in the flipper length column.

#### Scatterplot