---
title: "P8105 Homework #1"
author: "Zachary Katz (UNI: zak2132)"
date: "September 2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load tidyverse packages
library(tidyverse)
```

## Problem 1

Problem 1 asks us to create a data frame comprised of:

* A random sample of size 10 from a standard normal distribution
* A logical vector indicating whether elements of the sample are greater than 0
* A character vector of length 10
* A factor vector of length 10, with 3 different factor "levels"

### Create the data frame

First, we create the data frame by constructing four different variables (vectors), and then binding them together using the `tibble` command.

```{r create_data_frame}
# Set a seed to reproduce
set.seed(123)

# Create the variables/vectors separately

# numeric vector
rand_sample <- rnorm(10)
# logical vector if rand_sample is positive
logical_vector <- rand_sample > 0
# string vector of first 10 alphabet letters
character_vector <- letters[1:10]
# factor variable with 3 levels
factor_vector <- factor(sample(
      x = c("first", "second", "third"),
      size = 10,
      replace = TRUE))

# Construct a data frame (tibble) from the vectors
test_data_frame <- tibble(rand_sample, logical_vector, character_vector, factor_vector)

# Examine the data frame
test_data_frame

# Show the data frame's structure
str(test_data_frame)
```
By examining the structure of the data frame, we confirmed that the first column is a numeric variable, the second is a logical vector, the third is a character vector, and the fourth is a factor vector. Each vector has `r nrow(test_data_frame)` observations.

#### Find the means

Now that we've created the data frame, we try to take the mean of each variable.

```{r calculate_mean}
# Extract the columns one at a time and take a mean of each vector
sample_mean <- mean(pull(test_data_frame, rand_sample))
mean_positive <- mean(pull(test_data_frame, logical_vector))
mean_character_vector <- mean(pull(test_data_frame, character_vector))
mean_factor_vector <- mean(pull(test_data_frame, factor_vector))
```

* The mean of the random sample is `r sample_mean` and the mean of the logical vector is `r mean_positive`, which is the proportion of the random sample (of 10 observations) greater than 0.

* When we try to find the mean of the character and factor vectors, we learn that the mean function does not accept arguments that are not numeric or logical, and therefore returns NA.

#### Convert vectors

Finally, we want to convert the logical, character, and factor variables to numeric variables if possible.

```{r convert_to_numeric, results="hide"}
# Convert logical vector to numeric vector
convert_logical_to_numeric <- as.numeric(pull(test_data_frame, logical_vector))
convert_logical_to_numeric

# Convert character vector to numeric vector
convert_character_to_numeric <- as.numeric(pull(test_data_frame, character_vector))
convert_character_to_numeric

# Convert factor vector to numeric vector
convert_factor_to_numeric <- as.numeric(pull(test_data_frame, factor_vector))
convert_factor_to_numeric
```
* Converting the logical vector to a numeric vector changed **true to 1** and **false to 0**. 

* Converting the factor vector to a numeric vector changed **first level to 1, second level to 2, and third level to 3**.

* Converting the character vector to a numeric vector resulted in **"NAs introduced by coercion"**, leading to a vector composed only of 10 NAs.

## Problem 2

#### Load the penguins dataset

```{r load_penguins_dataset}
# Load penguins dataset
data("penguins", package = "palmerpenguins")
```

#### Describe the penguins dataset

First, we want to take a quick look at the `penguins` data and its structure.

```{r examine_penguins_data}
# Examine penguins data set
penguins

# Review head and tail of the data (first and last 10 rows, respectively)
head(penguins)
tail(penguins)

# Review structure of the data
str(penguins)

# Provide summary of the data
summary(penguins)

# Try another neat tool to show data distribution
skimr::skim(penguins)
```

The names of the variables in the dataset are `r names(penguins)` and the size of the dataset is `r nrow(penguins)` rows and `r ncol(penguins)` columns.

We are also asked to calculate the mean flipper length, which is `r mean(penguins$flipper_length_mm, na.rm = T)` mm, after removing NA entries in the flipper length column.

#### Create a scatterplot

Finally, we'd like to create and export a scatterplot of bill length vs. flipper length.

```{r scatterplot, warning=FALSE}

# Plot the graph
ggplot(data = penguins, 
       aes(x = bill_length_mm, 
           y = flipper_length_mm, 
           color = species)) + 
      geom_point() + 
   labs(
      title = "Flipper length vs. bill length",
      x = "Bill length (mm)",
      y = "Flipper length (mm)"
   )

# Save the graph
ggsave("flipper_v_bill_scatterplot.png")
```

This graph shows a positive linear relationship between penguin flipper length and bill length. This relationship seems to hold for each of the three species shown in the graph (Adelie, Chinstrap, Gentoo), though the trend for the Adelie species in particular seems to exhibit more variation.